### 2020-01-03 manuscript.Rmd
updates

---
title: "Notebook with AMP-AD RNA-seq network analysis results"
output: html_notebook
---

Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at synXXXver.x

```{r}
synapser::synLogin()

```

### 2020-01-03 R/buildAggregateModules.R

@@ -11,7 +11,6 @@ buildAggregateModules <- function(fullManifestOfDAta){
 
   #replace with indivudal module manifest
   #indMods <- fullManifestOfDAta$IndividualModules
-  synapser::synLogin()
   indMods <- synapser::synTableQuery("select * from syn10309369")$asDataFrame()
   indMods <- indMods[,-c(1,2)]
   #re-run enrichment analysis of full module set for these lists
@@ -149,5 +148,5 @@ buildAggregateModules <- function(fullManifestOfDAta){
                                    fp_mods$df)
 
   save(dlpfc_mods,cbe_mods,tcx_mods,ifg_mods,stg_mods,phg_mods,fp_mods,file='aggregateModules.rda')
-  rSynapseUtilities::makeTable(AggregateModuleManifest,'AMP-AD aggregate modules',projectId='syn2370594')
+  #rSynapseUtilities::makeTable(AggregateModuleManifest,'AMP-AD aggregate modules',projectId='syn2370594')
 }

### 2020-01-03 R/convertHgncToEnsembl.R
add utility functions to package

@@ -1,0 +1,13 @@
+convertHgncToEnsembl <- function(hgncIds){
+
+  library(biomaRt)
+  ensembl=biomaRt::useMart('ENSEMBL_MART_ENSEMBL',
+                           dataset = 'hsapiens_gene_ensembl',
+                           host='uswest.ensembl.org')
+
+  genes<-getBM(attributes = c('ensembl_gene_id','external_gene_name'),
+               filters='external_gene_name',
+               values=hgncIds,
+               mart=ensembl)
+  return(genes)
+}

### 2020-01-03 R/fisherWrapperOR.R

@@ -1,0 +1,10 @@
+fisherWrapperOR <- function(moduleGenes,annotationGenes,allGenes){
+  a00 <- sum(!(allGenes%in%moduleGenes) & !(allGenes%in%annotationGenes))
+  a10 <- sum((allGenes%in%moduleGenes) & !(allGenes%in%annotationGenes))
+  a01 <- sum(!(allGenes%in%moduleGenes) & (allGenes%in%annotationGenes))
+  a11 <- sum((allGenes%in%moduleGenes) & (allGenes%in%annotationGenes))
+  bar <- matrix(c(a00,a10,a01,a11),2,2)
+  #print(bar)
+  foo <- fisher.test(bar,alternative='greater')
+  return(as.numeric(foo$estimate)[1])
+}

### 2020-01-03 R/fisherWrapperPval.R

@@ -1,0 +1,10 @@
+fisherWrapperPval <- function(moduleGenes,annotationGenes,allGenes){
+  a00 <- sum(!(allGenes%in%moduleGenes) & !(allGenes%in%annotationGenes))
+  a10 <- sum((allGenes%in%moduleGenes) & !(allGenes%in%annotationGenes))
+  a01 <- sum(!(allGenes%in%moduleGenes) & (allGenes%in%annotationGenes))
+  a11 <- sum((allGenes%in%moduleGenes) & (allGenes%in%annotationGenes))
+  bar <- matrix(c(a00,a10,a01,a11),2,2)
+  #print(bar)
+  foo <- fisher.test(bar,alternative='greater')
+  return(foo$p.value)
+}

### 2020-01-03 R/outerSapply.R

@@ -1,0 +1,9 @@
+outerSapply <- function(FUN,X,Y,...){
+  require(dplyr)
+  pb <- progress::progress_bar$new(total=length(X))
+  internal <- function(X,Y,FUN,...){
+    pb$tick()
+    return(Y%>% sapply(FUN,X,...))
+  }
+  return(X %>% sapply(internal,Y,FUN,...))
+}

### 2020-01-03 R/removeSwappedDupKeyValueDf.R

@@ -1,0 +1,6 @@
+removeSwappedDupKeyValueDf <- function(x){
+  foo <- igraph::graph_from_data_frame(x,directed=FALSE)
+  foo <- igraph::simplify(foo,edge.attr.comb = list("first"))
+  foo <- igraph::as_data_frame(foo,what='edges')
+  return(foo)
+}

### 2020-01-03 manuscript.Rmd
remove utility function dependencies

@@
 
-Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at synXXXver.x
+Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at syn11932957
 

### 2020-01-03 R/buildAggregateModules.R

@@ -16,13 +16,13 @@ buildAggregateModules <- function(fullManifestOfDAta){
   #re-run enrichment analysis of full module set for these lists
 
   indModsList <- lapply(unique(indMods$ModuleNameFull),
-                        utilityFunctions::listify,
+                        AMPAD::listify,
                         indMods$GeneID,
                         indMods$ModuleNameFull)
   names(indModsList) <- unique(indMods$ModuleNameFull)
   allgenes <- unique(indMods$GeneID)
 
-  enrichmentMatrix <- utilityFunctions::outerSapply( utilityFunctions::fisherWrapperPval,
+  enrichmentMatrix <- AMPAD::outerSapply( AMPAD::fisherWrapperPval,
                                                      indModsList,
                                                      all.gs,
                                                      allgenes)
@@ -47,7 +47,7 @@ buildAggregateModules <- function(fullManifestOfDAta){
   pairMods <- synapser::synTableQuery("SELECT * FROM syn10339153 where ModuleNameFull like '%DLPFC' and category like '%DLPFC'")$asDataFrame()
   pairMods <- pairMods[,-c(1,2)]
   library(dplyr)
-  pairMods <- pairMods %>% utilityFunctions::removeSwappedDupKeyValueDf() %>%
+  pairMods <- pairMods %>% AMPAD::removeSwappedDupKeyValueDf() %>%
     dplyr::mutate(adj=p.adjust(fisherPval,method='bonferroni')) %>%
     dplyr::filter(adj<=0.05) %>%
     dplyr::filter(from%in%enrMatDLPFC$ModuleNameFull & to %in% enrMatDLPFC$ModuleNameFull) %>%
@@ -60,7 +60,7 @@ buildAggregateModules <- function(fullManifestOfDAta){
   pairMods <- synapser::synTableQuery("SELECT * FROM syn10339153 where ModuleNameFull like '%TCX' and category like '%TCX'")$asDataFrame()
   pairMods <- pairMods[,-c(1,2)]
   library(dplyr)
-  pairMods <- pairMods %>% utilityFunctions::removeSwappedDupKeyValueDf() %>%
[…]

### 2020-01-03 R/buildAggregateNonADModules.R

@@ -17,13 +17,13 @@ buildAggregateNonADModules <- function(fullManifestOfDAta){
   #re-run enrichment analysis of full module set for these lists
 
   indModsList <- lapply(unique(indMods$ModuleNameFull),
-                        utilityFunctions::listify,
+                        AMPAD::listify,
                         indMods$GeneID,
                         indMods$ModuleNameFull)
   names(indModsList) <- unique(indMods$ModuleNameFull)
   allgenes <- unique(indMods$GeneID)
 
-  enrichmentMatrix <- utilityFunctions::outerSapply( utilityFunctions::fisherWrapperPval,
+  enrichmentMatrix <- AMPAD::outerSapply( AMPAD::fisherWrapperPval,
                                                      indModsList,
                                                      all.gs,
                                                      allgenes)
@@ -48,7 +48,7 @@ buildAggregateNonADModules <- function(fullManifestOfDAta){
   pairMods <- synapser::synTableQuery("SELECT * FROM syn10339153 where ModuleNameFull like '%DLPFC' and category like '%DLPFC'")$asDataFrame()
   pairMods <- pairMods[,-c(1,2)]
   library(dplyr)
-  pairMods <- pairMods %>% utilityFunctions::removeSwappedDupKeyValueDf() %>%
+  pairMods <- pairMods %>% AMPAD::removeSwappedDupKeyValueDf() %>%
     dplyr::mutate(adj=p.adjust(fisherPval,method='bonferroni')) %>%
     dplyr::filter(adj<=0.05) %>%
     dplyr::filter(from%in%enrMatDLPFC$ModuleNameFull & to %in% enrMatDLPFC$ModuleNameFull) %>%
@@ -61,7 +61,7 @@ buildAggregateNonADModules <- function(fullManifestOfDAta){
   pairMods <- synapser::synTableQuery("SELECT * FROM syn10339153 where ModuleNameFull like '%TCX' and category like '%TCX'")$asDataFrame()
   pairMods <- pairMods[,-c(1,2)]
   library(dplyr)
[…]

### 2020-01-03 R/buildModuleGraph.R

@@ -72,7 +72,7 @@ buildModuleGraph <- function(pairwise,allMods,tissueType){
                  MoreArgs = list(tissueType=tissueType),
                  SIMPLIFY=F)
   res$df <- do.call(rbind,res$df)
-  exg <- utilityFunctions::convertEnsemblToHgnc(res$df$GeneID)
+  exg <- AMPAD::convertEnsemblToHgnc(res$df$GeneID)
   if(sum(duplicated(exg))>0){
     exg <- exg[which(!duplicated(exg)),]
   }

### 2020-01-03 R/collateEnrichmentSets.R

@@ -42,15 +42,15 @@ collateEnrichmentSets = function(){
 
   mckenzie <- data.table::fread('zhang_modules.csv',data.table = F)
   adList$MSSM <- dplyr::filter(mckenzie,Module == 'Red' | Module == 'List green' | Module == 'Green')$Gene_Symbol
-  enrichmentSets$MSSM2 <- utilityFunctions::listifyWrapper(mckenzie$Module,mckenzie$Gene_Symbol)
+  enrichmentSets$MSSM2 <- AMPAD::listifyWrapper(mckenzie$Module,mckenzie$Gene_Symbol)
 
 
   sczNetworkObj <- synapser::synGet('syn7118804')
   sczNetwork <- data.table::fread(sczNetworkObj$path,data.table=F)
-  enrichmentSets$scz <- utilityFunctions::listifyWrapper(sczNetwork$ModuleColor,sczNetwork$Ensembl)
+  enrichmentSets$scz <- AMPAD::listifyWrapper(sczNetwork$ModuleColor,sczNetwork$Ensembl)
   controlNetworkObj <- synapser::synGet('syn7118802')
   controlNetwork <- data.table::fread(controlNetworkObj$path,data.table=F)
-  enrichmentSets$control <- utilityFunctions::listifyWrapper(controlNetwork$ModuleColor,controlNetwork$Ensembl)
+  enrichmentSets$control <- AMPAD::listifyWrapper(controlNetwork$ModuleColor,controlNetwork$Ensembl)
 
 
   #load allen oligo data
@@ -67,7 +67,7 @@ collateEnrichmentSets = function(){
   allen_comprehensive$module <- paste0(allen_comprehensive$module,
                                        '_comprehensive')
   allen_list_s <- sapply(unique(allen_simple$module),
-                         utilityFunctions::listify,
+                         AMPAD::listify,
                          allen_simple$gene_symbol,
                          allen_simple$module)
   names(allen_list_s) <- unique(allen_simple$module)
@@ -75,7 +75,7 @@ collateEnrichmentSets = function(){
 
 
   allen_list_c <- sapply(unique(allen_comprehensive$module),
-                         utilityFunctions::listify,
+                         AMPAD::listify,
                          allen_comprehensive$gene_symbol,
                          allen_comprehensive$module)
 
@@ -103,7 +103,7 @@ collateEnrichmentSets = function(){
[…]

### 2020-01-03 R/improvedAdRelevancePlot.R

@@ -13,30 +13,30 @@ improvedAdRelevancePlot <- function(){
   foo2 <- synapser::synTableQuery("select distinct external_gene_name from syn10309369")$asDataFrame()
   foo2ensg <- synapser::synTableQuery("select distinct GeneID from syn10309369")$asDataFrame()
 
-  adListensg <- lapply(adList,utilityFunctions::convertHgncToEnsembl)
+  adListensg <- lapply(adList,AMPAD::convertHgncToEnsembl)
   adListensg2 <- lapply(adListensg,function(x) unique(x$ensembl_gene_id))
 
-  foobar <- utilityFunctions::outerSapplyParallel( utilityFunctions::fisherWrapperPval, amp.ad.de.geneSets, adListensg2,foo2ensg$GeneID)
+  foobar <- AMPAD::outerSapplyParallel( AMPAD::fisherWrapperPval, amp.ad.de.geneSets, adListensg2,foo2ensg$GeneID)
 
   foobar <- data.frame(foobar,stringsAsFactors=F)
   foobar$pathway <- rownames(foobar)
   foobar2<-tidyr::gather(foobar,key='geneset',value='pval',-pathway)
 
-  foobar3 <- utilityFunctions::outerSapplyParallel( utilityFunctions::fisherWrapperOR, amp.ad.de.geneSets, adListensg2,foo2ensg$GeneID)
+  foobar3 <- AMPAD::outerSapplyParallel( AMPAD::fisherWrapperOR, amp.ad.de.geneSets, adListensg2,foo2ensg$GeneID)
   foobar3 <- data.frame(foobar3,stringsAsFactors=F)
   foobar3$pathway <- rownames(foobar)
   foobar4<-tidyr::gather(foobar3,key='geneset',value='OR',-pathway)
 
   gaiteri_mods <- read.csv('zhang_modules.csv',stringsAsFactors=F)
-  modList <- lapply(unique(gaiteri_mods$Module), utilityFunctions::listify,gaiteri_mods$Gene_Symbol,gaiteri_mods$Module)
+  modList <- lapply(unique(gaiteri_mods$Module), AMPAD::listify,gaiteri_mods$Gene_Symbol,gaiteri_mods$Module)
   names(modList) <- unique(gaiteri_mods$Module)
[…]

### 2020-01-03 R/listifyWrapper.R

@@ -1,0 +1,9 @@
+listifyWrapper = function(x,y){
+  uniqueKeys <- unique(x)
+  res <- lapply(uniqueKeys,
+                AMPAD::listify,
+                y,
+                x)
+  names(res) <- uniqueKeys
+  return(res)
+}

### 2020-01-03 R/pairwiseMatrixOfEnrichments.R

@@ -8,7 +8,7 @@ pairwiseMatrixOfEnrichments = function(synId){
 
   #View(aggregateModules)
 
-  mats <- utilityFunctions::pairwiseMatrixOfEnrichments(key=aggregateModules$Module,
+  mats <- AMPAD::computePairwiseMatrixOfEnrichments(key=aggregateModules$Module,
                                                         value=aggregateModules$GeneID)
 
 

### 2020-01-03 R/run_amp_ad_enrichment.R

@@ -53,7 +53,7 @@ run_amp_ad_enrichment <- function(geneSetList,
   cat('running enrichments....\n')
 
   res <- list()
-  res$fisher <- utilityFunctions::outerSapplyParallel(utilityFunctions::fisherWrapper,
+  res$fisher <- AMPAD::outerSapplyParallel(AMPAD::fisherWrapper,
                                                       modulesLargeList,
                                                       geneSetList,
                                                       refGeneSet)

### 2020-01-03 R/run_amp_ad_enrichment2.R

@@ -53,7 +53,7 @@ run_amp_ad_enrichment2 <- function(geneSetList,
   cat('running enrichments....\n')
 
   res <- list()
-  res$fisher <- utilityFunctions::outerSapplyParallel(utilityFunctions::fisherWrapper,
+  res$fisher <- AMPAD::outerSapplyParallel(AMPAD::fisherWrapper,
                                                       modulesLargeList,
                                                       geneSetList,
                                                       refGeneSet)
@@ -67,7 +67,7 @@ run_amp_ad_enrichment2 <- function(geneSetList,
     return(length(intersect(x,y)))
   }
 
-  res$inter <- utilityFunctions::outerSapplyParallel(sizeOfInter,
+  res$inter <- AMPAD::outerSapplyParallel(sizeOfInter,
                                                      modulesLargeList,
                                                      geneSetList)
 

### 2020-01-03 R/run_amp_ad_enrichment2_lists.R

@@ -13,12 +13,12 @@ run_amp_ad_enrichment2_lists <- function(geneSetList,
   #OUTPUT:
   #data frame with module name, gene set name, p-value, and odds ratio of enrichment, intersections, and gene set sizes
   straightHgncConversion <- function(x){
-    y<-utilityFunctions::convertHgncToEnsembl(x)
+    y<-AMPAD::convertHgncToEnsembl(x)
     return(unique(y$ensembl_gene_id))
   }
 
   straightEnsemblConversion <- function(x){
-    y<- utilityFunctions::convertEnsemblToHgnc(x)
+    y<- AMPAD::convertEnsemblToHgnc(x)
     return(unique(y$external_gene_name))
   }
 
@@ -45,7 +45,7 @@ run_amp_ad_enrichment2_lists <- function(geneSetList,
   cat('running enrichments....\n')
 
   res <- list()
-  res$fisher <- utilityFunctions::outerSapplyParallel(utilityFunctions::fisherWrapper,
+  res$fisher <- AMPAD::outerSapplyParallel(AMPAD::fisherWrapper,
                                                       refGeneSetList,
                                                       geneSetList,
                                                       refGeneSet)
@@ -59,7 +59,7 @@ run_amp_ad_enrichment2_lists <- function(geneSetList,
     return(length(intersect(x,y)))
   }
 
-  res$inter <- utilityFunctions::outerSapplyParallel(sizeOfInter,
+  res$inter <- AMPAD::outerSapplyParallel(sizeOfInter,
                                                      refGeneSetList,
                                                      geneSetList)
 

### 2020-01-03 R/buildAggregateModules.R
non-ad specific module analysis

@@ -1,4 +1,4 @@
-buildAggregateModules <- function(fullManifestOfDAta){
+buildAggregateModules <- function(){
   #syn11914811
   #replace this with the appropriate synapse call soon
   fob1 <- synapser::synGet('syn11914811')
@@ -148,5 +148,6 @@ buildAggregateModules <- function(fullManifestOfDAta){
                                    fp_mods$df)
 
   save(dlpfc_mods,cbe_mods,tcx_mods,ifg_mods,stg_mods,phg_mods,fp_mods,file='aggregateModules.rda')
+  return(AggregateModuleManifest)
   #rSynapseUtilities::makeTable(AggregateModuleManifest,'AMP-AD aggregate modules',projectId='syn2370594')
 }

### 2020-01-03 R/buildAggregateNonADModules.R

@@ -1,4 +1,4 @@
-buildAggregateNonADModules <- function(fullManifestOfDAta){
+buildAggregateNonADModules <- function(){
   #syn11914811
   #replace this with the appropriate synapse call soon
   fob1 <- synapser::synGet('syn11914811')
@@ -11,7 +11,6 @@ buildAggregateNonADModules <- function(fullManifestOfDAta){
 
   #replace with indivudal module manifest
   #indMods <- fullManifestOfDAta$IndividualModules
-  synapser::synLogin()
   indMods <- synapser::synTableQuery("select * from syn10309369")$asDataFrame()
   indMods <- indMods[,-c(1,2)]
   #re-run enrichment analysis of full module set for these lists
@@ -147,7 +146,7 @@ buildAggregateNonADModules <- function(fullManifestOfDAta){
                                    stg_mods$df,
                                    phg_mods$df,
                                    fp_mods$df)
-
-  save(dlpfc_mods,cbe_mods,tcx_mods,ifg_mods,stg_mods,phg_mods,fp_mods,file='aggregateNonAdModules.rda')
+  return(AggregateModuleManifest)
+  #save(dlpfc_mods,cbe_mods,tcx_mods,ifg_mods,stg_mods,phg_mods,fp_mods,file='aggregateNonAdModules.rda')
   #rSynapseUtilities::makeTable(AggregateModuleManifest,'AMP-AD aggregate modules',projectId='syn2370594')
 }

### 2020-01-03 manuscript.Rmd
compare modules

@@
 
-Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at syn11932957
+Build aggregate modules from individual modules and differential expression meta-analysis, this will give you the same modules as are stored at syn11932957 - this code block builds the aggregate modules, and then checks if they are different from the paper aggregate modules.
 
@@
 synapser::synLogin()
+aggModsNew <- AMPAD::buildAggregateModules()
+aggModsPaper <- synapser::synTableQuery('select * from syn11932957')$asDataFrame()[,-c(1,2)]
+
+sum(aggModsNew[,1]!=aggModsPaper[,1])
+sum(aggModsNew[,2]!=aggModsPaper[,2])
+```
+
+Make the pairwise enrichment plot (Figure 1B)
+
+```{r}
+synapser::synLogin()
+AMPAD::pairwiseMatrixOfEnrichments('syn11932957')
+```
+
+Build aggregate modules without restricting the aggregation operation to the individual modules that are enriched for differentially expressed genes in AD.
+
+```{r}
+synapser::synLogin()
+aggModsNonAdNew <- AMPAD::buildAggregateNonADModules()
 
 ```
+
+Show overlap between AD and non AD (Figure SXF)
+```{r}
+AMPAD::compareADtoNonADmods(aggModsNew,
+                            aggModsNonAdNew)
+
+```

### 2020-01-03 R/compareADtoNonADmods.R

@@ -1,0 +1,38 @@
+compareADtoNonADmods <- function(ADaggMods,NonADaggMods,outputFile=FALSE){
+
+  foobar <- ADaggMods
+
+  AggregateModuleManifest <- NonADaggMods
+
+  mod1=c(paste0('all',unique(AggregateModuleManifest$Module)),paste0('ad',unique(foobar$Module)))
+  annrow <- data.frame(moduleType=c(rep('All',length(unique(AggregateModuleManifest$Module))),
+                                    rep('AD',length(unique(foobar$Module)))),
+                       stringsAsFactors=F)
+  rownames(annrow) <- mod1
+
+
+  mats <- utilityFunctions::pairwiseMatrixOfEnrichments(key=c(paste0('all',AggregateModuleManifest$Module),paste0('ad',foobar$Module)),
+                                                        value=c(AggregateModuleManifest$GeneID,foobar$GeneID))
+  pheatmap::pheatmap(-log10(mats$pval+1e-300),annotation_row=annrow)
+
+  if(outputFile){
+    tiff(filename='adnonadoverlap.tiff',height=85,width=85,units='mm',res=300)
+    pheatmap::pheatmap(-log10(mats$pval+1e-300),
+                       show_colnames = F,
+                       border_color = NA,
+                       fontsize=3,
+                       treeheight_row=10,
+                       treeheight_col=10,
+                       annotation_col=annrow)
+    dev.off()
+  }else{
+    pheatmap::pheatmap(-log10(mats$pval+1e-300),
+                       show_colnames = F,
+                       border_color = NA,
+                       fontsize=3,
+                       treeheight_row=10,
+                       treeheight_col=10,
+                       annotation_col=annrow)
+  }
+
+}

### 2020-01-03 R/pairwiseMatrixOfEnrichments.R

@@ -1,5 +1,5 @@
-pairwiseMatrixOfEnrichments = function(synId){
-  synapser::synLogin()
+pairwiseMatrixOfEnrichments = function(synId,outputFile=FALSE){
+
 
   ##pull aggregate modules
   #aggregateModules <- rSynapseUtilities::loadFullTable(synId)
@@ -53,19 +53,35 @@ pairwiseMatrixOfEnrichments = function(synId){
   #get rid of underflow issues
   #red,yellow,green,blue,purple
   mats$pval <- mats$pval + 1e-300
-  tiff(filename='figure2A.tiff',height=85,width=85,units='mm',res=300)
-  ann_colors<-list(Cluster=c(`Consensus Cluster A`='#F8766D',
-                 `Consensus Cluster B`='#A3A500',
-                 `Consensus Cluster C`='#00BF7D',
-                 `Consensus Cluster D`='#00B0F6',
-                 `Consensus Cluster E`='#E76BF3'))
-  pheatmap::pheatmap(-log10(mats$pval),
-                     show_colnames = F,
-                     border_color = NA,
-                     fontsize=4,
-                     treeheight_row=10,
-                     treeheight_col=10,
-                     annotation_col = customDf,
-                     annotation_colors = ann_colors[1])
-  dev.off()
+  if(outputFile){
+    tiff(filename='figure2A.tiff',height=85,width=85,units='mm',res=300)
+    ann_colors<-list(Cluster=c(`Consensus Cluster A`='#F8766D',
+                               `Consensus Cluster B`='#A3A500',
+                               `Consensus Cluster C`='#00BF7D',
+                               `Consensus Cluster D`='#00B0F6',
+                               `Consensus Cluster E`='#E76BF3'))
+    pheatmap::pheatmap(-log10(mats$pval),
+                       show_colnames = F,
+                       border_color = NA,
+                       fontsize=4,
+                       treeheight_row=10,
+                       treeheight_col=10,
+                       annotation_col = customDf,
+                       annotation_colors = ann_colors[1])
+    dev.off()
+  } else{
+    ann_colors<-list(Cluster=c(`Consensus Cluster A`='#F8766D',
+                               `Consensus Cluster B`='#A3A500',
+                               `Consensus Cluster C`='#00BF7D',
+                               `Consensus Cluster D`='#00B0F6',
+                               `Consensus Cluster E`='#E76BF3'))
+    pheatmap::pheatmap(-log10(mats$pval),
+                       show_colnames = F,
[…]

### 2020-01-03 R/collateEnrichmentSets.R
remove local dependencies

@@ -39,8 +39,8 @@ collateEnrichmentSets = function(){
   adList$genecards <- genecards$`Gene Symbol`
 
   #load mckenzie oligo data
-
-  mckenzie <- data.table::fread('zhang_modules.csv',data.table = F)
+  mckenzieObj1 <- synapser::synGet('syn21482836')
+  mckenzie <- data.table::fread(mckenzieObj1$path,data.table = F)
   adList$MSSM <- dplyr::filter(mckenzie,Module == 'Red' | Module == 'List green' | Module == 'Green')$Gene_Symbol
   enrichmentSets$MSSM2 <- AMPAD::listifyWrapper(mckenzie$Module,mckenzie$Gene_Symbol)
 
@@ -59,8 +59,12 @@ collateEnrichmentSets = function(){
 
   #adList$`AMP-AD Allen et al. 2018 AD+PSP TCX10.CS` <- dplyr::filter(allen,module=='TCX10')$gene_symbol
 
-  allen_simple <- data.table::fread('tcx_simple_oligo.csv',data.table=F)
-  allen_comprehensive <- data.table::fread('tcx_comprehensive_oligo.csv',data.table=F)
+
+  allen_simple_obj1 <- synapser::synGet('syn21482838')
+  allen_simple <- data.table::fread(allen_simple_obj1$path,data.table=F)
+
+  allen_comprehensive_obj1 <- synapser::synGet('syn21482837')
+  allen_comprehensive <- data.table::fread(allen_comprehensive_obj1$path,data.table=F)
 
   allen_simple$module <- paste0(allen_simple$module,
                                 '_simple')
@@ -93,7 +97,8 @@ collateEnrichmentSets = function(){
 
 
   #load johnson 2018
-  johnson <- data.table::fread('johnson.csv',data.table=F)
+  johnsonObj1 <- synapser::synGet('syn21482834')
+  johnson <- data.table::fread(johnsonObj1$path,data.table=F)
   adList$Emory <- dplyr::filter(johnson,Module == WGCNA::labels2colors(17) | Module == WGCNA::labels2colors(18) | Module == WGCNA::labels2colors(10) | Module == WGCNA::labels2colors(15) | Module == WGCNA::labels2colors(29) | Module == WGCNA::labels2colors(40))$geneName
   enrichmentSets$ad <- lapply(adList,unique)
 
[…]

### 2020-01-03 manuscript.Rmd
some updates

@@
 ```
+
+test building of gene sets
+
+```{r}
+targetedGeneSets <- AMPAD::collateEnrichmentSets()
+```
+
+Plot from Figure 1c
+```{r}
+synapser::synLogin()
+modMeta <- AMPAD::getModuleMetainfo('syn11932957')
+dummyDf <- targetedEnrichment$cell
+dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
+#dummyDf <- dplyr::filter(targetedEnrichment$cell,adj.pval<=0.05)
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+dummyDf$category <- gsub('Zhang\\.','',dummyDf$category)
+dummyDf$category <- factor(dummyDf$category,levels = rev(c('Astrocyte',
+                                                       'Endothelial',
+                                                       'Microglia',
+                                                       'Neuron',
+                                                       'MyelinOligos',
+                                                       'NewOligos',
+                                                       'OPC')))
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Cell Type Signature')
+g
+```

### 2020-01-03 R/make_dot_plot.R

@@ -1,0 +1,93 @@
+make_dot_plot <- function(dummyDf,xlab,ylab,cowsize=11,outputFile=FALSE,fileName='test.tiff'){
+  customDf <- data.frame(moduleName=c('TCXblue',
+                                      'IFGyellow',
+                                      'PHGyellow',
+                                      'DLPFCblue',
+                                      'CBEturquoise',
+                                      'STGblue',
+                                      'PHGturquoise',
+                                      'IFGturquoise',
+                                      'TCXturquoise',
+                                      'FPturquoise',
+                                      'IFGbrown',
+                                      'STGbrown',
+                                      'DLPFCyellow',
+                                      'TCXgreen',
+                                      'FPyellow',
+                                      'CBEyellow',
+                                      'PHGbrown',
+                                      'DLPFCbrown',
+                                      'STGyellow',
+                                      'PHGgreen',
+                                      'CBEbrown',
+                                      'TCXyellow',
+                                      'IFGblue',
+                                      'FPblue',
+                                      'FPbrown',
+                                      'CBEblue',
+                                      'DLPFCturquoise',
+                                      'TCXbrown',
+                                      'STGturquoise',
+                                      'PHGblue'),
+                         Cluster= c(rep('Consensus Cluster A',3),
+                                    rep('Consensus Cluster B',7),
+                                    rep('Consensus Cluster C',7),
+                                    rep('Consensus Cluster D',7),
+                                    rep('Consensus Cluster E',6)),
+                         stringsAsFactors=F)
+  dummyDf <- dplyr::left_join(dummyDf,customDf,by=c('Module'='moduleName'))
+  dummyDf$Cluster <- factor(dummyDf$Cluster,levels = (c('Consensus Cluster A',
+                                                        'Consensus Cluster B',
+                                                        'Consensus Cluster C',
+                                                        'Consensus Cluster D',
+                                                        'Consensus Cluster E')))
+  dummyDf$Module <- factor(dummyDf$Module,levels = (c('TCXblue',
+                                                      'IFGyellow',
+                                                      'PHGyellow',
+                                                      'DLPFCblue',
+                                                      'CBEturquoise',
+                                                      'STGblue',
+                                                      'PHGturquoise',
+                                                      'IFGturquoise',
+                                                      'TCXturquoise',
+                                                      'FPturquoise',
+                                                      'IFGbrown',
+                                                      'STGbrown',
+                                                      'DLPFCyellow',
[…]

### 2020-01-03 R/run_full_enrichment_suite.R

@@ -1,0 +1,81 @@
+run_full_enrichment_suite <- function(){
+  targetedGeneSets <- AMPAD::collateEnrichmentSets()
+
+  targetedEnrichment <- list()
+  targetedEnrichment$ad <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$ad,
+                                                         'AD',
+                                                         hgnc = TRUE,
+                                                         manifestId = 'syn11932957')
+
+  targetedEnrichment$cell <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$cell,
+                                                           'Cell',
+                                                           hgnc= TRUE,
+                                                           manifestId = 'syn11932957')
+
+  targetedEnrichment$cell2 <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$cell2,
+                                                            'Mathys',
+                                                            hgnc= TRUE,
+                                                            manifestId = 'syn11932957')
+
+  targetedEnrichment$scz <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$scz,
+                                                          'scz',
+                                                          hgnc= FALSE,
+                                                          manifestId = 'syn11932957')
+
+  targetedEnrichment$control <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$control,
+                                                              'control',
+                                                              hgnc= FALSE,
+                                                              manifestId = 'syn11932957')
+
+  targetedEnrichment$mssm2 <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$MSSM2,
+                                                            'zhang',
+                                                            hgnc= TRUE,
+                                                            manifestId = 'syn11932957')
+
+  targetedEnrichment$mito <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$mito,
+                                                           'Mito',
+                                                           hgnc= TRUE,
+                                                           manifestId = 'syn11932957')
+
+  targetedEnrichment$deg <- AMPAD::run_amp_ad_enrichment2(targetedGeneSets$deg,
+                                                          'DEG',
+                                                          hgnc=FALSE,
[…]

### 2020-01-03 manuscript.Rmd
add plots

@@
 
-Show overlap between AD and non AD (Figure SXF)
+Show overlap between AD and non AD (Figure S2A)
 ```{r}
@@
 ```{r}
-targetedGeneSets <- AMPAD::collateEnrichmentSets()
+targetedEnrichment <- AMPAD::run_full_enrichment_suite()
 ```
@@
 ```
+
+Plot from Figure S2B
+
+```{r}
+dummyDf <- targetedEnrichment$cell2
+dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
+#dummyDf <- dplyr::filter(targetedEnrichment$cell,adj.pval<=0.05)
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Lake et al. Cell Type Signature')
+g
+```
+
+Plot for figure S6B
+
+```{r}
+dummyDf <- targetedEnrichment$scz
+dummyDf <- dplyr::filter(dummyDf,category == 'yellow' | category == 'red' | category == 'blue' | category == 'lightyellow' | category == 'greenyellow' | category == 'cyan' | category == 'grey60')
+dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
+#dummyDf <- dplyr::filter(targetedEnrichment$cell,adj.pval<=0.05)
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='CommonMind Differentially Expressed Module')
+g
+```
+
+plot from figure S6C
+
+```{r}
+dummyDf <- targetedEnrichment$mssm2
+dummyDf <- dplyr::filter(dummyDf,category == 'Yellow' | category == 'Pink' | category == 'Gray.1' | category == 'Seashell' | category == 'Red.3' | category == 'Green.yellow' | category == 'Red' | category == 'Gold.2' | category == 'Tan' | category == 'Gold.3' | category == 'Light.yellow' | category == 'Brown.2' | category == 'Dark.cyan' | category == 'Khaki' | category == 'Grey.60' | category == 'Purple' | category == 'Green.4' | category == 'Honey.dew' | category == 'Red.2' | category == 'Beige')
+dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
+#dummyDf <- dplyr::filter(targetedEnrichment$cell,adj.pval<=0.05)
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Zhang et al. Module')
+g
+```
+
+Plot from figure 1D
+
+```{r}
+dummyDf <- targetedEnrichment$ad
+dummyDf$fisherOR[dummyDf$adj.pval>=0.05] <- NA
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+#dummyDf$ModuleBrainRegion <- factor(dummyDf$ModuleBrainRegion,levels = unique(dummyDf$ModuleBrainRegion))
+dummyDf$category <- factor(dummyDf$category,
+                           levels = rev(c('genecards',
+                                       'pantherPresenilin',
+                                       'dbgap',
+                                       'igap',
+                                       'jensenDisease',
+                                       'omimExpanded',
+                                       'biocarta',
+                                       'wikipathwaysMouse',
+                                       'wikipathwaysHuman',
+                                       'pantherAmyloid',
+                                       'kegg',                                       
+                                       'omim',
+                                       'Nominated_targets',
+                                       'MSSM',
+                                       'Mayo_simple',
+                                       'Mayo_comprehensive',
+                                       'Emory',
+                                       'Columbia_Broad_Rush_m109')))
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='AD Signature')
+g
+
+```
+
+plot from figure 4D
+
+```{r}
+dummyDf <- dplyr::filter(targetedEnrichment$degMeta,adj.pval<=0.05)
+dummyDf <- dplyr::filter(dummyDf,category == 'ad.control.FEMALE.random.DOWN' |
+                           category == 'ad.control.FEMALE.random.UP' | 
+                           category == 'ad.control.MALE.random.DOWN' | 
+                           category == 'ad.control.MALE.random.UP')
+dummyDf$adj.pval[dummyDf$adj.pval==0] = 10^-300
+dummyDf$category <- gsub('ad\\.control\\.FEMALE\\.random\\.DOWN','Female, Down',dummyDf$category)
+dummyDf$category <- gsub('ad\\.control\\.MALE\\.random\\.DOWN','Male, Down',dummyDf$category)
+dummyDf$category <- gsub('ad\\.control\\.FEMALE\\.random\\.UP','Female, Up',dummyDf$category)
+dummyDf$category <- gsub('ad\\.control\\.MALE\\.random\\.UP','Male, Up',dummyDf$category)
+dummyDf <- dplyr::left_join(dummyDf,modMeta)
+
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='DEG Sex Specific Meta Analysis')
+g
+
+```

### 2020-01-03 R/run_full_enrichment_suite.R

@@ -78,4 +78,5 @@ run_full_enrichment_suite <- function(){
   targetedEnrichment$targetPathways$adj.pval <- p.adjust(targetedEnrichment$targetPathways$fisherPval,method='fdr')
   targetedEnrichment$degFull$adj.pval <- p.adjust(targetedEnrichment$degFull$fisherPval,method='fdr')
   targetedEnrichment$degMetaFull$adj.pval <- p.adjust(targetedEnrichment$degMetaFull$fisherPval,method='fdr')
+  return(targetedEnrichment)
 }

### 2020-01-03 manuscript.Rmd
magma and upset updates

@@
 
+Show that the aggregate modules are better enriched than other potential AD modules (Figure 1A)
+
+```{r}
+synapser::synLogin()
+g <- AMPAD::improvedAdRelevancePlot()
+g
+```
+
+Show summaries of overlaps with UpSet plots (Figure S1A-S1E)
+
+```{r}
+synapser::synLogin()
+res<-AMPAD::aggregate_module_summary_plots()
+res
+```
+
 Build aggregate modules without restricting the aggregation operation to the individual modules that are enriched for differentially expressed genes in AD.
@@
 targetedEnrichment <- AMPAD::run_full_enrichment_suite()
+```
+
+Run MAGMA analysis (must have magma tool installed in current working directory that this notebook is running from), producing the plot from Figure S1F
+```{r}
+synapser::synLogin()
+g <- AMPAD::run_magma_analysis()
+g
 ```

### 2020-01-03 R/aggregate_module_summary_plots.R

@@ -1,0 +1,139 @@
+aggregate_module_summary_plots = function(outputFile=FALSE){
+  foo <- synapser::synTableQuery("select * from syn11932957")$asDataFrame()
+  foo2 <- dplyr::select(foo,GeneID,Module)
+  foo2$Presence <- 1
+  foo3 <- tidyr::pivot_wider(foo2,
+                             id_cols = "GeneID",
+                             names_from = "Module",
+                             values_from = "Presence")
+  foo3[is.na(foo3)] <- 0
+  foo3 <- data.frame(foo3,stringsAsFactors=F)
+  foo4 <- dplyr::select(foo3,GeneID,TCXblue,IFGyellow,PHGyellow)
+  resu <- list()
+
+  if(outputFile){
+    tiff(filename = 'consensusClusterA.tiff', height = 4, width = 6,units='in',pointsize=14,res=300)
+    UpSetR::upset(foo4,nintersects = NA,show.numbers=F)
+    dev.off()
+  }else{
+    resu$A<-UpSetR::upset(foo4,nintersects = NA,show.numbers=F)
+  }
+
+  nUniqueGenesA <- data.frame(Module=c('TCXblue','PHGyellow','IFGyellow'),nGenes=c(979,366,127),stringsAsFactors=F)
+
+  foo4 <- dplyr::select(foo3,GeneID,DLPFCblue,CBEturquoise,STGblue,PHGturquoise,IFGturquoise,TCXturquoise,FPturquoise)
+
+
+  if(outputFile){
+    tiff(filename = 'consensusClusterB.tiff', height = 4, width = 6,units='in',pointsize=14,res=300)
+    UpSetR::upset(foo4,nsets=7,nintersects = NA,point.size=1,show.numbers = F)
+    dev.off()
+  } else{
+    resu$B<-UpSetR::upset(foo4,nsets=7,nintersects = NA,point.size=1,show.numbers = F)
+  }
+
+  nUniqueGenesB <- data.frame(Module=c('CBEturquoise','DLPFCblue','IFGturquoise','PHGturquoise','STGblue','TCXturquoise','FPturquoise'),nGenes=c(593,349,275,209,163,69,40),stringsAsFactors=F)
+
+  foo4 <- dplyr::select(foo3,GeneID,IFGbrown,STGbrown,DLPFCyellow,TCXgreen,FPyellow,CBEyellow,PHGbrown)
+
+  if(outputFile){
[…]

### 2020-01-03 R/improvedAdRelevancePlot.R

@@ -1,5 +1,5 @@
-improvedAdRelevancePlot <- function(){
-  synapser::synLogin()
+improvedAdRelevancePlot <- function(outputFile=FALSE){
+
   adGeneticsSummaryAggOld <- AMPAD::getAdGenetics(synId='syn10915669')
   adGeneticsSummaryAgg <- AMPAD::getAdGenetics(synId='syn11926100')
   adGeneticsSummaryInd <- AMPAD::getAdGenetics(synId='syn10309369')
@@ -27,7 +27,10 @@ improvedAdRelevancePlot <- function(){
   foobar3$pathway <- rownames(foobar)
   foobar4<-tidyr::gather(foobar3,key='geneset',value='OR',-pathway)
 
-  gaiteri_mods <- read.csv('zhang_modules.csv',stringsAsFactors=F)
+
+  mckenzieObj1 <- synapser::synGet('syn21482836')
+  gaiteri_mods <- data.table::fread(mckenzieObj1$path,data.table = F)
+  #gaiteri_mods <- read.csv('zhang_modules.csv',stringsAsFactors=F)
   modList <- lapply(unique(gaiteri_mods$Module), AMPAD::listify,gaiteri_mods$Gene_Symbol,gaiteri_mods$Module)
   names(modList) <- unique(gaiteri_mods$Module)
   foobarX <- AMPAD::outerSapplyParallel( AMPAD::fisherWrapperPval, modList, adList,foo2$external_gene_name)
@@ -59,7 +62,9 @@ improvedAdRelevancePlot <- function(){
   gaiteriEnrichment <- dplyr::left_join(foobarX2,foobarX4)
   degAnalysisEnrichment <- dplyr::left_join(foobar2,foobar4)
 
-  load('aggregateModules.rda')
+  #load('aggregateModules.rda')
+  aggmodobj <- synapser::synGet('syn21483261')
+  load(aggmodobj$path)
   combinedMatrix <- rbind(dlpfc_mods$moduleGraph,
                           cbe_mods$moduleGraph,
                           tcx_mods$moduleGraph,
@@ -192,8 +197,12 @@ improvedAdRelevancePlot <- function(){
 
 
   #
-  g
+  #g
   #g2 <- cowplot::plot_grid(g)
-  ggplot2::ggsave('figure1.tiff',device='tiff',units='mm',width=114,height=85,scale=2)
+  if(outputFile){
+    g
[…]

### 2020-01-03 R/make_dot_plot.R

@@ -87,6 +87,7 @@ make_dot_plot <- function(dummyDf,xlab,ylab,cowsize=11,outputFile=FALSE,fileName
   g <- g + AMPAD::cowplot_rotated(cowsize)
 
   if(outputFile){
+    g
     ggplot2::ggsave(fileName,device='tiff',units='mm',width=85,height=85,scale=1.8)
   }
   return(g)

### 2020-01-03 R/run_magma_analysis.R

@@ -1,0 +1,169 @@
+run_magma_analysis = function(outputFile=FALSE){
+
+
+  #download stage 2 data
+  str <- "curl -O https://www.niagads.org/system/tdf/public_docs/Kunkle_etal_Stage2_results.txt?file=1&type=field_collection_item&id=121&force="
+
+  system(str)
+
+
+  #rename file
+  str <- "mv Kunkle_etal_Stage2_results.txt\\?file\\=1 Kunkle_etal_Stage2_results.txt"
+  system(str)
+
+  #load into R, and munge into appropriate format
+  pvals <- data.table::fread('Kunkle_etal_Stage2_results.txt')
+
+  #SNP,CHR,BP,P,NOBS
+  pvalsReformatted <- dplyr::select(pvals,
+                                    MarkerName,
+                                    Chromosome,
+                                    Position,
+                                    Pvalue)
+
+  colnames(pvalsReformatted) <- c('SNP',
+                                  'CHR',
+                                  'BP',
+                                  'P')
+
+  pvalsReformatted$NOBS <- 82771
+  pvalsReformatted$P <- as.numeric(pvalsReformatted$P)
+
+  write.table(pvalsReformatted,sep="\t",row.names=F,file="kunkle_stage2.txt",quote=F)
+
+  write.table(pvalsReformatted[,1:3],sep="\t",row.names=F,col.names=F,file="kunkle2snp.txt",quote=F)
+
+  str <- ".//magma --annotate --snp-loc kunkle2snp.txt --gene-loc NCBI37.3.gene.loc --out kunkleannots2"
+
+
+  system(str)
+
+  str <- ".//magma --bfile g1000_eur --pval kunkle_stage2.txt N=82771 --gene-annot kunkleannots2.genes.annot --out kunkle2 --gene-model multi=snp-wise"
+  system(str)
+
+
+  #create geneset files
+  #pull manifest from synapse
+
+  synapser::synLogin()
+  ampMods <- synapser::synTableQuery("select * from syn11932957")$asDataFrame()
+
+
+  #convert ensembl ids to ncbi ids
+
+  convertEnsemblToEntrez <- function(ensemblIds){
+
+    library(biomaRt)
+    ensembl=biomaRt::useMart('ENSEMBL_MART_ENSEMBL',
+                             dataset = 'hsapiens_gene_ensembl',
+                             host='uswest.ensembl.org')
+
[…]

### 2020-01-03 manuscript.Rmd
updated scripts

@@
 g <- AMPAD::run_magma_analysis()
+g
+```
+
+Run expression weighted cell enrichment method (must have downloaded and unzipped the flat file from GSE97930 in your current working directory: https://www.ncbi.nlm.nih.gov/geo/download/?acc=GSE97930&format=file&file=GSE97930%5FFrontalCortex%5FsnDrop%2Dseq%5FUMI%5FCount%5FMatrix%5F08%2D01%2D2017%2Etxt%2Egz )
+
+This step also may require a large compute instance (with > 64 Gb of memory) to successfully run.
+
+```{r}
+synapser::synLogin()
+g <- AMPAD::run_ewce()
 g

### 2020-01-03 archive/cross_study_sample_viz.Rmd

---
title: "R Notebook"
output: html_notebook
---

ROSMAP data present
```{r}
synapser::synLogin()
foo <- synapser::synTableQuery("select * from syn18912660")$asDataFrame()
foo <- dplyr::filter(foo,study=='ROSMAP')
bar <- dplyr::select(foo,individualID,dataType,assay,futureData)

fxn1 <- function(x,y){
  if(x=='metabolomics'){
    return(x)
  }else{
    return(y)
  }
}

bar$dataDescriptor <- mapply(fxn1,bar$dataType,bar$assay)
bar$exists <- bar$futureData
bar$exists[bar$exists==TRUE] <- 2
bar$exists[is.na(bar$exists)] <- 1
bar <- dplyr::select(bar,individualID,dataDescriptor,exists)
uniqueId <- paste0(bar$individualID,bar$dataDescriptor)
dup <- duplicated(uniqueId)
dupInd <- uniqueId[dup]
keepIndex <- uniqueId%in%dupInd

bar$exists[keepIndex] <- 3
bar <- bar[!dup,]

#bar <- bar[!dup,]
baz <- tidyr::spread(bar,key=individualID,value=exists)
baz[is.na(baz)] <- 0
rownames(baz) <- baz$dataDescriptor
baz <- baz[,-1]
baz2 <- t(apply(baz,1,as.numeric))

ab<-pheatmap::pheatmap(baz2)
baz3 <- baz2[ab$tree_row$order,ab$tree_col$order]
png(file='ROSMAP.png',width=2400,height=1600,pointsize=14)
pheatmap::pheatmap(baz3,color = rev(viridis::viridis(4)),cluster_rows = F,cluster_cols =F,legend_breaks=0:3,legend_labels=c('Absent\n\n\n\n','Available\n\n','Planned','\n\nAvailable\n+Planned'))
dev.off()
```

```{r}
synapser::synLogin()
foo <- synapser::synTableQuery("select * from syn18912660")$asDataFrame()
foo <- dplyr::filter(foo,study=='MayoRNAseq')
bar <- dplyr::select(foo,individualID,dataType,assay,futureData)

bar$dataDescriptor <- bar$assay
bar$exists <- bar$futureData
bar$exists[bar$exists==TRUE] <- 2
bar$exists[is.na(bar$exists)] <- 1
bar <- dplyr::select(bar,individualID,dataDescriptor,exists)
uniqueId <- paste0(bar$individualID,bar$dataDescriptor)
dup <- duplicated(uniqueId)
dupInd <- uniqueId[dup]
keepIndex <- uniqueId%in%dupInd

bar$exists[keepIndex] <- 3
bar <- bar[!dup,]

#bar <- bar[!dup,]
baz <- tidyr::spread(bar,key=individualID,value=exists)
baz[is.na(baz)] <- 0
rownames(baz) <- baz$dataDescriptor
baz <- baz[,-1]
baz2 <- t(apply(baz,1,as.numeric))

ab<-pheatmap::pheatmap(baz2)
baz3 <- baz2[ab$tree_row$order,ab$tree_col$order]
png(file='ROSMAP.png',width=2400,height=1600,pointsize=14)
pheatmap::pheatmap(baz3,color = rev(viridis::viridis(4)),cluster_rows = F,cluster_cols =F,legend_breaks=0:3,legend_labels=c('Absent\n\n\n\n','Available\n\n','Planned','\n\nAvailable\n+Planned'))
dev.off()
```

```{r}
synapser::synLogin()
foo <- synapser::synTableQuery("select * from syn18912660")$asDataFrame()
foo <- dplyr::filter(foo,study=='MSBB')
bar <- dplyr::select(foo,individualID,dataType,assay,futureData)

fxn1 <- function(x){
  if(is.na(x)){
    return('MSBBproteomics')
  }else{
    return(x)
  }
}

bar$dataDescriptor <- sapply(bar$assay,fxn1)
#bar$dataDescriptor <- bar$assay
bar$exists <- bar$futureData
bar$exists[bar$exists==TRUE] <- 2
bar$exists[is.na(bar$exists)] <- 1
bar <- dplyr::select(bar,individualID,dataDescriptor,exists)
uniqueId <- paste0(bar$individualID,bar$dataDescriptor)
dup <- duplicated(uniqueId)
dupInd <- uniqueId[dup]
keepIndex <- uniqueId%in%dupInd

bar$exists[keepIndex] <- 3
bar <- bar[!dup,]

#bar <- bar[!dup,]
baz <- tidyr::spread(bar,key=individualID,value=exists)
baz[is.na(baz)] <- 0
rownames(baz) <- baz$dataDescriptor
baz <- baz[,-1]
baz2 <- t(apply(baz,1,as.numeric))

ab<-pheatmap::pheatmap(baz2)
baz3 <- baz2[ab$tree_row$order,ab$tree_col$order]
png(file='ROSMAP.png',width=2400,height=1600,pointsize=14)
pheatmap::pheatmap(baz3,color = rev(viridis::viridis(4)),cluster_rows = F,cluster_cols =F,legend_breaks=0:3,legend_labels=c('Absent\n\n\n\n','Available\n\n','Planned','\n\nAvailable\n+Planned'))
dev.off()
```

### 2020-01-12 manuscript.Rmd
some figure updates

@@
 AMPAD::pairwiseMatrixOfEnrichments('syn11932957')
+AMPAD::pairwiseMatrixOfEnrichments('syn11932957',
+                                   outputFile = T,
+                                   fileName1 = '~/Desktop/updatedFiguresJan72020/figure1b.tiff')
 ```
@@
 g
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Cell Type Signature',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figure1c.tiff')
 ```
@@
 g
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Lake et al. Cell Type Signature',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figureS2b.tiff')
 ```
@@
 g
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='CommonMind Differentially Expressed Module',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figures6b.tiff')
+
 ```
@@
 g
+
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='Zhang et al. Module',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figures6c.tiff')
 ```
@@
 #dummyDf$ModuleBrainRegion <- factor(dummyDf$ModuleBrainRegion,levels = unique(dummyDf$ModuleBrainRegion))
+dummyDf <- dplyr::filter(dummyDf,category!='Nominated_targets' & category!='Mayo_simple' & category != 'omim')
+
+dummyDf$category <- gsub('Mayo_comprehensive',
+                         'Mayo_RNAseq',
+                         dummyDf$category)
+# dummyDf$category <- factor(dummyDf$category,
+#                            levels = rev(c('genecards',
+#                                        'pantherPresenilin',
+#                                        'dbgap',
+#                                        'igap',
+#                                        'jensenDisease',
+#                                        'omimExpanded',
+#                                        'biocarta',
+#                                        'wikipathwaysMouse',
+#                                        'wikipathwaysHuman',
+#                                        'pantherAmyloid',
+#                                        'kegg',                                       
+#                                        'omim',
+#                                        'Nominated_targets',
+#                                        'MSSM',
+#                                        'Mayo_simple',
+#                                        'Mayo_comprehensive',
+#                                        'Emory',
+#                                        'Columbia_Broad_Rush_m109')))
+
 dummyDf$category <- factor(dummyDf$category,
@@
                                        'kegg',                                       
-                                       'omim',
-                                       'Nominated_targets',
                                        'MSSM',
-                                       'Mayo_simple',
-                                       'Mayo_comprehensive',
+                                       'Mayo_RNAseq',
                                        'Emory',
                                        'Columbia_Broad_Rush_m109')))
+
 g<-AMPAD::make_dot_plot(dummyDf,
@@
 
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='AD Signature',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figure1d.tiff',
+                 scaleValue = 2)
 ```
@@
 
+g<-AMPAD::make_dot_plot(dummyDf,
+                 xlab='AD Coexpression Module',
+                 ylab='DEG Sex Specific Meta Analysis',
+                 outputFile = TRUE,
+                 fileName = '~/Desktop/updatedFiguresJan72020/figure4d.tiff')
 ```

### 2020-01-12 R/collateEnrichmentSets.R

@@ -23,15 +23,15 @@ collateEnrichmentSets = function(){
   adList <- adList[-which(adList=='PSEN1' | adList =='PSEN2' | adList =='APP' | adList == 'P2K2B')]
   adList <- list(igap = adList)
   adList$dbgap <- dbgap$`Alzheimer Disease`
-  adList$kegg <- kegg$`Alzheimer's disease_Homo sapiens_hsa05010`
-  adList$wikipathwaysMouse <- wikipathways$`Alzheimers Disease_Mus musculus_WP2075`
-  adList$wikipathwaysHuman <- wikipathways$`Alzheimers Disease_Homo sapiens_WP2059`
-  adList$jensenDisease <- jensen_diseases$`Alzheimer's_disease`
-  adList$biocarta <- biocarta$`Deregulation of CDK5 in Alzheimers Disease_Homo sapiens_h_p35alzheimersPathway`
-  adList$pantherAmyloid <- panther$`Alzheimer disease-amyloid secretase pathway_Homo sapiens_P00003`
-  adList$pantherPresenilin <- panther$`Alzheimer disease-presenilin pathway_Homo sapiens_P00004`
-  adList$omim <- omim_disease$`alzheimer_disease`
-  adList$omimExpanded <- omim_expanded$`alzheimer_disease`
+  adList$kegg <- kegg$`Alzheimer's disease Homo sapiens hsa05010`
+  adList$wikipathwaysMouse <- wikipathways$`Alzheimers Disease Mus musculus WP2075`
+  adList$wikipathwaysHuman <- wikipathways$`Alzheimers Disease Homo sapiens WP2059`
+  adList$jensenDisease <- jensen_diseases$`Alzheimer's disease`
+  adList$biocarta <- biocarta$`Deregulation of CDK5 in Alzheimers Disease Homo sapiens h p35alzheimersPathway`
+  adList$pantherAmyloid <- panther$`Alzheimer disease-amyloid secretase pathway Homo sapiens P00003`
+  adList$pantherPresenilin <- panther$`Alzheimer disease-presenilin pathway Homo sapiens P00004`
+  adList$omim <- omim_disease$`alzheimer disease`
+  adList$omimExpanded <- omim_expanded$`alzheimer disease`
 
   genecardsObj <- synapser::synGet('syn10507702')
   genecards <- data.table::fread(genecardsObj$path,data.table=F)

### 2020-01-12 R/make_dot_plot.R

@@ -1,4 +1,4 @@
-make_dot_plot <- function(dummyDf,xlab,ylab,cowsize=11,outputFile=FALSE,fileName='test.tiff'){
+make_dot_plot <- function(dummyDf,xlab,ylab,cowsize=11,outputFile=FALSE,fileName='test.tiff',scaleValue=1.8){
   customDf <- data.frame(moduleName=c('TCXblue',
                                       'IFGyellow',
                                       'PHGyellow',
@@ -83,12 +83,17 @@ make_dot_plot <- function(dummyDf,xlab,ylab,cowsize=11,outputFile=FALSE,fileName
 
   g <- g + ggplot2::labs(y = ylab,
                          x = xlab)
-
   g <- g + AMPAD::cowplot_rotated(cowsize)
+
+  g <- g + ggplot2::scale_color_manual(values = c('#fefd11',
+                                                  '#18bebf',
+                                                  '#a82828',
+                                                  '#34cc37',
+                                                  '#470606'))
 
   if(outputFile){
     g
-    ggplot2::ggsave(fileName,device='tiff',units='mm',width=85,height=85,scale=1.8)
+    ggplot2::ggsave(fileName,device='tiff',units='mm',width=85,height=85,scale=scaleValue)
   }
   return(g)
 }

### 2020-01-12 R/run_ewce.R

@@ -149,6 +149,11 @@ run_ewce <- function(outputFile=FALSE){
   #g <- g + ggplot2::ggtitle('Enrichment for Cell Type Specific Signatures')
   g <- g + ggplot2::labs(y = 'Lake et al. Cell Type Signature',
                          x = 'AD Coexpression Module')
+  g <- g + ggplot2::scale_color_manual(values = c('#fefd11',
+                                                  '#18bebf',
+                                                  '#a82828',
+                                                  '#34cc37',
+                                                  '#470606'))
   g <- g + AMPAD::cowplot_rotated(11)
 
 

### 2020-01-12 R/getAdGenetics2.R
updates

@@ -51,7 +51,9 @@ getAdGenetics2 <- function(synId='syn10338156'){
   adList <- adList[-which(adList=='HLA-DRB5-DRB1')]
   adList <- adList[-which(adList=='PSEN1' | adList =='PSEN2' | adList =='APP' | adList == 'P2K2B')]
 
+
   adList <- list(igap = adList)
+
   adList$dbgap <- dbgap$`Alzheimer Disease`
   adList$kegg <- kegg$`Alzheimer's disease_Homo sapiens_hsa05010`
   adList$wikipathwaysMouse <- wikipathways$`Alzheimers Disease_Mus musculus_WP2075`

### 2020-02-05 manuscript.Rmd
updates

@@
 #dummyDf$ModuleBrainRegion <- factor(dummyDf$ModuleBrainRegion,levels = unique(dummyDf$ModuleBrainRegion))
-dummyDf <- dplyr::filter(dummyDf,category!='Nominated_targets' & category!='Mayo_simple' & category != 'omim')
+dummyDf <- dplyr::filter(dummyDf,category!='Nominated_targets' & category!='Mayo_simple' & category != 'omim' & category != 'jensenDisease' & category != 'wikipathwaysMouse' & category != 'pantherAmyloid')
 
 dummyDf$category <- gsub('Mayo_comprehensive',
-                         'Mayo_RNAseq',
+                         'Mayo Oligo Modules',
                          dummyDf$category)
+
+dummyDf$category <- gsub('Columbia_Broad_Rush_m109',
+                         'ROSMAP m109',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('Emory',
+                         'Emory AD Modules',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('MSSM',
+                         'MSSM Oligo Modules',
+                         dummyDf$category)
+
 # dummyDf$category <- factor(dummyDf$category,
@@
                                        'igap',
-                                       'jensenDisease',
                                        'omimExpanded',
                                        'biocarta',
-                                       'wikipathwaysMouse',
                                        'wikipathwaysHuman',
-                                       'pantherAmyloid',
                                        'kegg',                                       
@@
                  outputFile = TRUE,
-                 fileName = '~/Desktop/updatedFiguresJan72020/figure1d.tiff',
+                 fileName = '~/Desktop/updatedFiguresFeb52020/figure1d.tiff',
                  scaleValue = 2)

### 2020-02-05 R/collateEnrichmentSets.R

@@ -41,7 +41,7 @@ collateEnrichmentSets = function(){
   #load mckenzie oligo data
   mckenzieObj1 <- synapser::synGet('syn21482836')
   mckenzie <- data.table::fread(mckenzieObj1$path,data.table = F)
-  adList$MSSM <- dplyr::filter(mckenzie,Module == 'Red' | Module == 'List green' | Module == 'Green')$Gene_Symbol
+  adList$MSSM <- dplyr::filter(mckenzie,Module == 'Red' | Module == 'Light green' | Module == 'Green')$Gene_Symbol
   enrichmentSets$MSSM2 <- AMPAD::listifyWrapper(mckenzie$Module,mckenzie$Gene_Symbol)
 
 

### 2020-09-22 manuscript.Rmd
updates

@@
 #dummyDf$ModuleBrainRegion <- factor(dummyDf$ModuleBrainRegion,levels = unique(dummyDf$ModuleBrainRegion))
+
 dummyDf <- dplyr::filter(dummyDf,category!='Nominated_targets' & category!='Mayo_simple' & category != 'omim' & category != 'jensenDisease' & category != 'wikipathwaysMouse' & category != 'pantherAmyloid')
 
-dummyDf$category <- gsub('Mayo_comprehensive',
+dummyDf$category <- gsub('Mayo\\_comprehensive',
                          'Mayo Oligo Modules',
@@
 
-dummyDf$category <- gsub('Columbia_Broad_Rush_m109',
+dummyDf$category <- gsub('Columbia\\_Broad\\_Rush\\_m109',
                          'ROSMAP m109',
@@
                          'MSSM Oligo Modules',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('igap',
+                         'IGAP',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('biocarta',
+                         'Biocarta',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('genecards',
+                         'Genecards',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('kegg',
+                         'KEGG',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('wikipathwaysHuman',
+                         'Wikipathways Human',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('pantherPresenilin',
+                         'Panther Presenilin',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('dbgap',
+                         'dbGaP',
+                         dummyDf$category)
+
+dummyDf$category <- gsub('omimExpanded',
+                         'OMIM Expanded',
                          dummyDf$category)
@@
 dummyDf$category <- factor(dummyDf$category,
-                           levels = rev(c('genecards',
-                                       'pantherPresenilin',
-                                       'dbgap',
-                                       'igap',
-                                       'omimExpanded',
-                                       'biocarta',
-                                       'wikipathwaysHuman',
-                                       'kegg',                                       
-                                       'MSSM',
-                                       'Mayo_RNAseq',
-                                       'Emory',
-                                       'Columbia_Broad_Rush_m109')))
+                           levels = rev(c('Genecards',
+                                       'Panther Presenilin',
+                                       'dbGaP',
+                                       'IGAP',
+                                       'OMIM Expanded',
+                                       'Biocarta',
+                                       'Wikipathways Human',
+                                       'KEGG',                                       
+                                       'MSSM Oligo Modules',
+                                       'Mayo Oligo Modules',
+                                       'Emory AD Modules',
+                                       'ROSMAP m109')))
 
@@
                  xlab='AD Coexpression Module',
-                 ylab='AD Signature',
+                 ylab='AD Gene Set',
                  outputFile = TRUE,
